---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# iai

<!-- badges: start -->
<!-- badges: end -->

The `iai` package implements methods to estimate conditional outcome means in settings with missingness-not-at-random and incomplete auxiliary variables. Specifically, this package implements the AF4 method in Mathur et al. (In preparation).

## Installation

You can install the development version of `iai` from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("stmcg/iai")
```

## Example

We first load the package.
```{r}
library(iai)
```

#### Data Set

We will use the example dataset `dat.sim` included the package. The dataset contains 9,297 observations with a continuous outcome `Y`, a binary auxiliary variable `W`, and binary predictors `X1` and `X2`. The first 10 rows of `dat.sim` are: 
```{r}
dat.sim[1:10,]
```

#### AF4 Method

The AF4 method estimates $E_{AF4}[ Y \mid X=x ]$, which is identified by $\int_{W} E [ Y | X=x, W, M=1 ] p( W | X=x, R_W = R_X = 1 ) dW$ where $R_W$ and $R_X$ are indicators of non-missing values of $W$ and $X$, respectively, and $M$ is an indicator of a complete case pattern (i.e., $Y$, $X$, and $W$ are non-missing). The AF4 method estimates $E_{AF4} \big[ Y \mid X=x \big]$ by fitting models for the conditional mean of $Y$ and conditional density of $W$ and performing Monte Carlo integration to compute the integral. 

The function `af4` implements the AF4 method. This function requires specifying the following models: 

* `Y_model`: Formula for the outcome model
* `W_model`: Formula for the auxiliary model

It also requires specifying the names of the variable(s) $X$ by `X_names` and their values $x$ in $E_{AF4} \big[ Y \mid X=x \big]$ by `X_values`. 

An application of `af4` to estimate $E_{AF4} \big[ Y \mid X_1=0, X_2 = 1 \big]$ is given below:

```{r}
set.seed(1234)
res <- af4(data = dat.sim,
           X_names = c("X1", "X2"), X_values = c(0, 1),
           Y_model = Y ~ W + X1 + X2, W_model = W ~ X1 + X2)
```

The estimated outcome mean is given below:
```{r}
res$mean_est
```
