---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# iai

<!-- badges: start -->
[![R-CMD-check](https://github.com/stmcg/iai/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/stmcg/iai/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/stmcg/iai/graph/badge.svg)](https://app.codecov.io/gh/stmcg/iai)
<!-- badges: end -->

The `iai` package implements methods to estimate conditional outcome means in settings with missingness-not-at-random and incomplete auxiliary variables. Specifically, this package implements the AF4 method in Mathur et al. (In preparation).

## Installation

You can install the development version of `iai` from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("stmcg/iai")
```

## Example

We first load the package.
```{r}
library(iai)
```

#### Data Set

We will use the example dataset `dat.sim` included in the package. The dataset contains 9,297 observations with a continuous outcome `Y`, a binary auxiliary variable `W`, and binary predictors `X1` and `X2`. The first 10 rows of `dat.sim` are: 
```{r}
dat.sim[1:10,]
```

#### AF4 Method

The AF4 method estimates $E_{AF4}[ Y | X=x ]$, which is identified by $\int_{W} E [ Y | X=x, W, M=1 ] p( W | X=x, R_W = R_X = 1 ) dW$ where $R_W$ and $R_X$ are indicators of non-missing values of $W$ and $X$, respectively, and $M$ is an indicator of a complete case pattern (i.e., $Y$, $X$, and $W$ are non-missing). The AF4 method estimates $E_{AF4}[ Y | X=x]$ by fitting models for the conditional mean of $Y$ and conditional density of $W$ and performing Monte Carlo integration to compute the integral. 

The function `af4` implements the AF4 method to estimate $E_{AF4}[ Y | X=x_1 ]$ and, optionally, $E_{AF4}[ Y | X=x_2]$ as well as contrasts between $E_{AF4}[ Y | X=x_1 ]$ and $E_{AF4}[ Y | X=x_2 ]$ (differences, ratios). This function requires specifying the following models: 

* `Y_model`: Formula for the outcome model
* `W_model`: Formula for the auxiliary model when the auxiliary variable is univariate, or a list of formulas for each component of the auxiliary variable

It also requires specifying the names of the variable(s) $X$ by `X_names` and their values $x_1$ and $x_2$ by `X_values_1` and `X_values_2`, respectively. 

An application of `af4` to estimate $E_{AF4} [ Y | X=(0, 1)^\top ]$ and $E_{AF4} [ Y | X = (0, 0)^\top ]$ as well as their difference is given below:

```{r}
set.seed(1234)
res <- af4(data = dat.sim,
           X_names = c("X1", "X2"), 
           X_values_1 = c(0, 1), X_values_2 = c(0, 0),
           Y_model = Y ~ W + X1 + X2, W_model = W ~ X1 + X2)
res
```

We can obtain 95\% confidence intervals around our estimates by applying the `get_CI` function to the output of the `af4` function. The `get_CI` function performs nonparametric bootstrap. Here, we use the percentile method with 100 bootstrap replicates (for ease of computation): 

```{r}
get_CI(res, n_boot = 100, type = 'perc')
```
